---
title: "Final Project "
author: "Allison Dong & Afshin Sardaryzadeh"
date: "5/21/2020"
output: html_document
---

## Data
For the final group project, we will use ALL of the data sets provided under this GitHub account (global confirmed, recovered, and deaths as well as US confirmed and deaths).

Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE). Data includes daily time series CSV summary tables, including confirmations and deaths. Lat and Long refer to coordinates references for the user. Date fields are stored in MM/DD/YYYY format.

## Background
The Global Health Initiative has recently employed a new data science response team, CSIT-165, that uses data science to characterize pandemic diseases. CSIT-165 disseminates data driven analyses to global and local decision makers.
CSIT-165 is a conglomerate comprised of two fabricated entities: World Health Organization (WHO) and U.S. Pandemic Response Team (USPRT). 

## Link to GitHub repository
https://github.com/allison-d-g/final-project

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(magrittr)
library(readr)
library(geosphere)
library(dplyr)
library(tidyr)
library(tibble)
library(tidyverse)
library(stringr)
library(leaflet)
library(httr)
library(ggplot2)
library(scales)
library(cowplot)
```

## Global trajectory

```{r gt}
options(scipen=999)
##create a key value to later join tables##
Confirmedwithkey <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")%>%
  unite("Key",Lat:Long,remove=FALSE)
deathswithkey <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")%>%
  unite("Key",Lat:Long,remove=FALSE)
recoveredwithkey <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")%>%
  unite("Key",Lat:Long,remove=FALSE)
##----------------------------##
##Add the status to each table##
Confirmedwithkey$Status<-'Confirmed'
deathswithkey$Status<-'Deaths'
recoveredwithkey$Status<-'Recovered'
##----------------------------##
##Join tables##
Combo1<-rbind(Confirmedwithkey,deathswithkey)
BigATable<-rbind(Combo1,recoveredwithkey)
##----------------------------##
##Manipulate tables##
BigATable2<-BigATable%>%
  gather(key=date,
         value=cases,
         dplyr::ends_with('0'))%>%
  group_by(Status)%>%
mutate(Weekdate=as.Date(date, format="%m/%d/%y"))
##----------------------------##
##Group and Summarize Table##
  BigATable3<-BigATable2%>%
    group_by(Status,Weekdate)%>%
  summarize(Totalcases = sum(cases))
##----------------------------##
##Create Graph##
ggplot(data=BigATable3)+geom_point(mapping=aes(x=Weekdate,y=Totalcases, color=Status, shape=Status)) +
labs(
    title ="Global Trajectory of COVID-19",
    color = "Global Data Set" ) 
##----------------------------##
```
## Global Map

```{r gm}
##Group and Summarize Table##
  BigATableMap<-BigATable2%>%
    group_by(Status,Lat,Long)%>%
  summarize(Totalcases = sum(cases))
##----------------------------##
##Create 3 tables to filter on##
BigATableMapDeaths<- BigATableMap%>% filter(Status=="Deaths")
BigATableMapConfirmed<- BigATableMap%>% filter(Status=="Confirmed")
BigATableMapRecovered<- BigATableMap%>% filter(Status=="Recovered")
##----------------------------##
##Create the interactive map##
leaflet(data = BigATableMap) %>% 
 addTiles() %>%
  addCircleMarkers(data=BigATableMapDeaths, ~Long , ~Lat, radius=8 , color="red",
                   fillColor="red", stroke = TRUE, fillOpacity = 0.8, group="Deaths") %>%
  addCircleMarkers(data=BigATableMapConfirmed, ~Long , ~Lat, radius=8 , color="green",
                   fillColor="green", stroke = TRUE, fillOpacity = 0.8, group="Recovered") %>%
    addCircleMarkers(data=BigATableMapRecovered, ~Long , ~Lat, radius=8 , color="orange",fillColor="orange", stroke = TRUE, fillOpacity = 0.8, group="Confirmed") %>%
  
  addLayersControl(
       overlayGroups = c("Recovered", "Deaths", "Confirmed"),
    options = layersControlOptions(collapsed = FALSE)
  )
##----------------------------##
```

## Narrowing Down Hotspots
```{r ndh}
##Group and Summarize Table##
  NiceTable<-BigATable2%>%
    group_by(Status,`Country/Region`)%>%
  summarize(Totalcases = sum(cases))
##----------------------------##
NiceTable%>%
  kable()%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"
))%>%
  kable_styling(fixed_thead = T)%>%
pack_rows(index=table(NiceTable$Status))
 #group_rows(index=NiceTable$Status)
#collapse_rows(columns = 1, valign = "top")
```

## Zooming Into Our State
```{r zios}
# Get the data from the repository
us_confirmed <-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")

us_death <-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")

 # Create a dataset with data for california
us_confirmed_ca <- us_confirmed %>% filter(str_detect(us_confirmed$Province_State, 'California'))
us_death_ca <- us_death %>% filter(str_detect(us_death$Province_State, 'California'))

# Delete the other unnecessary columns
us_confirmed_ca_con <- select(us_confirmed_ca, -UID, -iso2, -iso3, -code3, -FIPS, -Admin2, -Country_Region, -Combined_Key, -Lat, -Long_)
us_death_ca_con <- select(us_death_ca, -UID, -iso2, -iso3, -code3, -FIPS, -Admin2, -Country_Region, -Combined_Key, -Lat, -Long_, -Population)

# Sum the number of confirmed and death for california
us_confirmed_ca_grouped <- us_confirmed_ca_con %>%
  group_by(Province_State) %>%
  summarise_all(funs(sum))

us_death_ca_grouped <- us_death_ca_con %>%
  group_by(Province_State) %>%
  summarise_all(funs(sum))

# Change the date format to normal date format
us_confirmed_ca_grouped_rearrange <- 
   gather(us_confirmed_ca_grouped, key = Date, value = Cases, -Province_State)
us_confirmed_ca_grouped_rearrange <- data.frame(us_confirmed_ca_grouped_rearrange, stringsAsFactors = FALSE) %>%
  mutate(Date=as.Date(Date,format = "X%m.%d.%y")) 

us_death_ca_grouped_rearrange <- 
   gather(us_death_ca_grouped, key = Date, value = Cases, -Province_State)
us_death_ca_grouped_rearrange <- data.frame(us_death_ca_grouped_rearrange, stringsAsFactors = FALSE) %>%
  mutate(Date=as.Date(Date,format = "X%m.%d.%y")) 

# Plot the results for confirmed and deaths case in CA
plot1 <- ggplot() +
  geom_point(data = us_confirmed_ca_grouped_rearrange, aes(x = Date, y= Cases, color ="blue")) +
  geom_point(data = us_death_ca_grouped_rearrange, aes(x = Date, y= Cases, color = "red"))+
  geom_vline(xintercept = as.numeric(as.Date("2020-03-19")), color = "black") +
  labs(
    title ="California's Trajectory for COVID-19") +
  scale_colour_hue("California Total", labels = c("Confirmed", "Deaths")) 

# Delete the other unnecessary columns
us_confirmed_city_con <- select(us_confirmed_ca, -UID, -iso2, -iso3, -code3, -FIPS, -Province_State, -Country_Region, -Combined_Key, -Lat, -Long_)

# Rank the data by the city has highest cases 
confirmed_cases_city <- arrange(us_confirmed_city_con, -us_confirmed_city_con$X5.20.20)

# Get the top 3 cities
top_3_confirmed <- filter(confirmed_cases_city[1:3, ])

# Change the date format to normal date format
confirmed_cases_city_rearrange <- 
   gather(top_3_confirmed, key = Date, value = Cases, -Admin2)
confirmed_cases_city_rearrange <- data.frame(confirmed_cases_city_rearrange, stringsAsFactors = FALSE) %>%
  mutate(Date=as.Date(Date,format = "X%m.%d.%y")) 

# Plot the top 3 cities
plot2 <- ggplot(data = confirmed_cases_city_rearrange) +
  geom_point(
  mapping = aes(x = Date, y= Cases,color = Admin2)) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-19")), color = "black") +
  labs(
    title ="California's top three city trajectories for confirmations",
    color = "Top Confirmed Cities" ) 

plot_grid(plot1, plot2, ncol = 1, align = 'v')
```
## Digging Deeper
```{r DD}
# Delete the other unnecessary columns
confirmed_population <- select(us_confirmed, -UID, -iso2, -iso3, -code3, -FIPS, -Country_Region, -Combined_Key, -Lat, -Long_, -Province_State)

death <- select(us_death, -UID, -iso2, -iso3, -code3, -FIPS, -Country_Region, -Combined_Key, -Lat, -Long_, -Province_State, -Admin2)

# Change the date format to normal date format 
#and change from wide to long
confirmed_population_rearrange <- 
   gather(confirmed_population, key = Date, value = Confirmed, -Admin2)
confirmed_population_rearrange <- data.frame(confirmed_population_rearrange, stringsAsFactors = FALSE) %>%
  mutate(Date=as.Date(Date,format = "X%m.%d.%y"))

death_rearrange <- 
   gather(death, key = Date, value = Death, -Population)
death_rearrange <- data.frame(death_rearrange, stringsAsFactors = FALSE) %>%
  mutate(Date=as.Date(Date,format = "X%m.%d.%y"))

# Plot for population vs. confirmation
plot.population <- ggplot(data = data.frame(confirmed=confirmed_population_rearrange$Confirmed, Population=us_death$Population)) +
  geom_point(
  mapping = aes(x = Population, y= confirmed), color = "blue") +
  scale_x_continuous(labels=scales::comma, trans = log2_trans()) +
  scale_y_continuous(labels=scales::comma, trans = log2_trans()) +
  labs(
    title ="Confirmations Vs. Population", y = "Confirmation Counts") 

# Plot for confirmation vs. death
plot.con_death <- ggplot(data = data.frame(Confirmed=confirmed_population_rearrange$Confirmed, Death=death_rearrange$Death)) +
  geom_point(
  mapping = aes(x = Confirmed, y= Death), color = "red") +
  scale_x_continuous(labels=scales::comma, trans = log2_trans()) +   
  scale_y_continuous(labels=scales::comma, trans = log2_trans()) +
  labs(
    title ="Deaths VS. Confirmations", x= "Confirmed Counts", y="Deaths Counts") 
  
plot_grid(plot.population, plot.con_death, align = 'v')
```

```{bash}
git log --pretty=format:"%nSubject: %s%nAuthor: %aN%nDate: %aD%nBody: %b"
```
